Revisiting the hydrodynamic wave modulation from linear wave theory

### 2024-08-29 paper/paper.tex
Wavenumber modulation edits

\begin{abstract}
Hydrodynamic modulation of short ocean surface waves by longer ambient waves is
a well-known ocean surface process that affects remote sensing, the
interpretation of \textit{in situ} wave measurements, and numerical wave
forecasting.
In this paper, we revisit the linear wave theory and derive higher-order
solutions for the change of short-wave wavenumber, action density, and
gravitational acceleration due to the presence of longer waves.
We validate the analytical solutions with numerical simulations of the full wave
crest and action conservation equations.
The nonlinear analytical solutions of short-wave wavenumber, amplitude, and
steepness modulation significantly deviate from the linear analytical solutions
of \citet{longuet1960changes}, and are similar to the nonlinear numerical
solutions by \citet{zhang1990evolution}.
We also discuss the result of \citet{peureux2021unsteady} who found through
numerical simulations that the short-wave modulation grows unsteadily with
each long-wave phase.
Here we show that this unsteady growth only occurs for homogeneous initial
conditions and does not occur for more realistic long-wave groups.
Finally, the combined analytical and numerical framework allows us to quantify
the contribution of various terms in the full wave conservation equations.
The new analytical solutions are a good approximation of the full numerical
solution in long-wave steepness up to $\approx 0.2$.
\end{abstract}

@@

For the ambient forcing, we consider a train of monochromatic long wave
defined, to the first order in steepness, by the elevation $\eta_L = a_L cos{\psi}$,
where $a_L$ is the long-wave amplitude, $\psi = k_L x - \omega_L t$ is the
long-wave phase, and $\omega_L$ its angular frequency.
At this time we define the long-wave steepness $\varepsilon_L = a_L k_L$ which
we will use throughout the rest of the paper as a scaling parameter, as is common
in the literature.
The long wave induces horizontal and vertical orbital velocities whose values
at the surface $\eta_L$ are:

@@

We now describe the analytical solutions for the short-wave wavenumber,
effective gravitational acceleration, amplitude, steepness, intrinsic frequency,
and intrinsic phase speed in presence of long waves.

@@

\begin{equation}
\label{eq:k_short_exact}
\widetilde{k}(\psi) = k e^{\varepsilon_L \cos{\psi} e^{\varepsilon_L \cos{\psi}}}
\end{equation}

Notice that the Taylor expansion of Eq. \ref{eq:k_short_exact} to the first order
recovers the original solution by L-HS:

\begin{equation}
\label{eq:k_short_lhs}
\widetilde{k}(\psi) = k (1 + \varepsilon_L \cos{\psi})
\end{equation}

The L-HS solution requires two approximations in their solution:
First, evaluating the long-wave velocity at $z = 0$
rather than $z = \eta_L$; and second, evaluating the solution to the first order
of the Taylor expansion series.
Both of these approximations cause an underestimate of the short-wave
modulation magnitude (Fig. \ref{fig:wavenumber_modulation}).
The original L-HS solution is a cosine wave with an amplitude equal
to the long-wave steepness.
The other two solutions both introduce an increase in wave modulation
on the crest and a decrease in the trough.
At the crest, the relative modulation predicted by the exact L-HS solution
is 10.7\% larger than that predicted by the 1st order approximate L-HS
solution, while the solution that evaluates the long-wave orbital velocity
at the water surface rather than $z = 0$ is 38.4\% larger than that of the
1$^{st}$ order approximate L-HS solution.

>>>>

\begin{abstract}
Hydrodynamic modulation of short ocean surface waves by longer ambient waves is
a well-known ocean surface process that affects remote sensing, the
interpretation of \textit{in situ} wave measurements, and numerical wave
forecasting.
In this paper, we revisit the linear wave theory and derive higher-order
solutions for the change of short-wave wavenumber, action density, and
gravitational acceleration due to the presence of longer waves.
We validate the analytical solutions with numerical simulations of the full wave
crest and action conservation equations.
The nonlinear analytical solutions of short-wave wavenumber, amplitude, and
steepness modulation significantly deviate from the linear analytical solutions
of \citet{longuet1960changes}, and are similar to the nonlinear numerical
solutions by \citet{longuet1987propagation} and \citet{zhang1990evolution}.
We also examine the result of \citet{peureux2021unsteady} who found through
numerical simulations that the short-wave modulation grows unsteadily with
each long-wave phase.
We show that this unsteady growth only occurs for homogeneous initial
conditions and does not occur for more realistic long-wave groups.
The combined analytical and numerical framework presented here allows us to
quantify the importance of various terms in the full wave conservation
equations.
The new analytical solutions are a good approximation of the full numerical
solution in long-wave steepness up to $\approx 0.2$.
\end{abstract}

@@

For the ambient forcing, we consider a train of monochromatic long wave
defined, to the first order in steepness, by the elevation $\eta_L = a_L cos{\psi}$,
where $\psi = k_L x - \omega_L t$ is the long-wave phase and $\omega_L$ its
angular frequency.
The long wave induces horizontal and vertical orbital velocities whose values
at the surface $\eta_L$ are:

@@

We now describe the analytical solutions for the short-wave wavenumber,
effective gravitational acceleration, amplitude, steepness, intrinsic frequency,
and intrinsic phase speed in presence of long waves.
The wavenumber is derived by linearizing the conservation of wave crests
(Eq. \ref{eq:wave_crests}) and neglecting the spatial gradients of $k$.
The effective gravity is derived by projecting the long-wave induced surface
accelerations onto the axis that is normal to the long-wave surface.
(Eq. \ref{eq:wave_action2}) and using the expression for the long-wave orbital
velocities.
The amplitude is derived by first evaluation the short-wave action from the
action conservation equation (Eq. \ref{eq:wave_action}) and then applying the
modulated effective gravity to obtain the amplitude.
From the modulated wavenumber, amplitude, and effective gravity, the steepness,
frequency, and phase speed readily follow from the usual wave relationships.
Going forward we will use the tilde over the variables to denote the modulated
quantities (e.g. $\widetilde{k}$ for wavenumber).

@@

\begin{equation}
\label{eq:k_short_exact}
\widetilde{k} = k e^{\varepsilon_L \cos{\psi} e^{\varepsilon_L \cos{\psi}}}
\end{equation}

Notice that the Taylor expansion of Eq. \ref{eq:k_short_exact} to the first order
recovers the original solution by \citet{longuet1960changes}:

\begin{equation}
\label{eq:k_short_lhs}
\widetilde{k} = k (1 + \varepsilon_L \cos{\psi})
\end{equation}
Coming from the conservation of wave crests, as opposed to the velocity potential
superposition approach, the L-HS solution requires two approximations in their
solution:
First, evaluating the long-wave velocity at $z = 0$ rather than $z = \eta_L$; and
second, further truncating the Taylor expansion series beyond the first order term.
These approximations cause an underestimate of the short-wave modulation
magnitude (Fig. \ref{fig:analytical_solutions_ak0.1}a).
At the wave crest, Eq. \ref{eq:k_short_exact} yields the wavenumber modulation
that is 16.9\%, 38.5\%, 66.4\%, and 104\% larger than that of \citet{longuet1960changes}
for $\varepsilon_L = 0.1, 0.2, 0.3, 0.4$, respectively.
